---
title: "ToothGrowth Statistical Inference P2"
author: "Zamora, Juan Carlo"
date: "August 19, 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
### Overview
This project analyzes the ToothGrowth data set by comparing the guinea tooth growth by supplement and dose. 
Contains 1. Data analysis; and 2. Comparison with confidence intervals to make conclusions about the tooth growth.

### Load the ToothGrowth data and perform exploratory data analyses
```{r results='hide'}
library(datasets)
data(ToothGrowth)
str(ToothGrowth)
head(ToothGrowth)
summary(ToothGrowth)
```

```{r}
library(ggplot2)
t = ToothGrowth
levels(t$supp) <- c("Orange Juice", "Ascorbic Acid")
ggplot(t, aes(x=factor(dose), y=len)) + 
  facet_grid(.~supp) +
  geom_boxplot(aes(fill = supp), show_guide = FALSE) +
  labs(title="Guinea pig tooth length by dosage for each type of supplement", 
    x="Dose (mg/day)",
    y="Tooth Length")
```

### Basic summary of the data
The box plot shows that dosage and tooth length are directly proportional. Orange juice is more effective than ascorbic acid for tooth growth when the dosage is .5 to 1.0 milligrams per day. Both types of supplements are equally as effective when the dosage is 2.0 milligrams per day.

### Confidence Intervals & Hypothesis Tests
#### Hypothesis #1
Orange juice & ascorbic acid deliver the same tooth growth.
```{r}
hypoth1<-t.test(len ~ supp, data = t)
hypoth1$conf.int
hypoth1$p.value
```
The confidence intervals includes 0 and the p-value is greater than the 
threshold of 0.05.  Accept the null hypothesis.

#### Hypothesis #2
For the dosage of 0.5 mg/day, the two supplements deliver the similar growth.
```{r}
hypoth2<-t.test(len ~ supp, data = subset(t, dose == 0.5))
hypoth2$conf.int
hypoth2$p.value
```
The confidence interval does not include 0 and the p-value is below the 0.05 threshold. Reject Null hypothesis.

#### Hypothesis #3
For the dosage of 1 mg/day, the two supplements deliver the similar growth
```{r}
hypoth3<-t.test(len ~ supp, data = subset(t, dose == 1))
hypoth3$conf.int
hypoth3$p.value
```
The confidence interval does not include 0 and the p-value is smaller than the 
0.05 threshold. The null hypothesis can be rejected. The alternative hypothesis 
that 1 mg/day dosage of orange juice delivers more tooth growth than ascorbic 
acid is accepted.

#### Hypothesis #4
For the dosage of 2 mg/day, the two supplements deliver the similar growth
```{r}
hypoth4<-t.test(len ~ supp, data = subset(t, dose == 2))
hypoth4$conf.int
hypoth4$p.value
```
The confidence interval does include 0 and the p-value is larger than the 
0.05 threshold. The null hypothesis cannot be rejected.


### Conclusions & Discussion
Orange juice delivers more tooth growth than ascorbic acid for lower dosages. Orange juice and ascorbic acid deliver the ssimilar growth for higher dose.  We cannot conclude orange juice is more effective that ascorbic acid. Recommending to perform other statistical methods for verifications

Assumptions

* Normal distribution of the tooth lengths
* No other unmeasured factors are affecting tooth length