---
title: "Statistical Inference P1 August"
author: "Zamora, Juan Carlo"
date: "August 19, 2018"
output:  
  html_document: default
  pdf_document: default
  word_document: default
---

## Overview
The purpose of this data analysis is to investigate the exponential distribution
and compare it to the Central Limit Theorem. For this analysis, the lambda will 
be set to 0.2 for all of the simulations. This investigation will compare the 
distribution of averages of 40 exponentials over 1000 simulations.


## Simulations
Set the simulation variables lambda, exponentials, and seed.
```{r}
ECHO=TRUE
set.seed(6969)
lambda = 0.2
exponentials = 40
```

Run Simulations with variables
```{r}
Means = NULL
for (i in 1 : 1000) Means = c(Means, mean(rexp(exponentials, lambda)))
```


## Sample Mean versus Theoretical Mean
#### Sample Mean
Calculating the mean from the simulations with give the sample mean.
```{r}
mean(Means)
```

#### Theoretical Mean
The theoretical mean of an exponential distribution is lambda^-1.
```{r}
lambda^-1
```

#### Difference of Means
Minimal difference from the simulation and the theoretical values.
```{r}
abs(mean(Means)-lambda^-1)
```


## Sample Variance versus Theoretical Variance
#### Sample Variance
Calculate the variance
```{r}
var(Means)
```

#### Theoretical Variance
The theoretical variance of an exponential distribution is 
(lambda * sqrt(n))^-2.
```{r}
(lambda * sqrt(exponentials))^-2
```

#### Difference of variance
Minimal difference from the simulation and the theoretical values.
```{r}
abs(var(Means)-(lambda * sqrt(exponentials))^-2)
```


## Distribution
This is a density histogram of the 1000 simulations. There is an overlay with a 
normal distribution that has a mean of lambda^-1 and standard deviation of 
(lambda*sqrt(n))^-1, the theoretical normal distribution for the simulations.
```{r}
library(ggplot2)
ggplot(data.frame(y=Means), aes(x=y)) + 
  geom_histogram(aes(y=..density..), binwidth=0.0625, fill="#E70F00", 
                 color="pink") +
  stat_function(fun=dnorm, args = list(mean=lambda^-1, 
                                    sd=(lambda*sqrt(exponentials))^-1), 
                size=1.5) +
  labs(title="Simulation", x="Mean")
```
